---
title: "FInal Exam"
author: "Prashant Gupta"
Vid: "V01038654"
output: html_document
---


```{r message=FALSE, warning=FALSE}
library(forecast)
library(ggplot2)
```

```{r}
data= read.csv("AustralianTourism.csv")
```

```{r}
head(data)
```


```{r}
data.ts <- ts(data$Victoria, start = c(1998, 1), end = c(2017, 4), freq = 4)
```

### Using 1998 to 2014 as the training set, keeping 2015, 2016, and 2017 as the validation set.

```{r}
nValid <- 12
nTrain <- length(data.ts) - nValid

# partition the data
train.ts <- window(data.ts, start = c(1998, 1), end = c(1998, nTrain))
valid.ts <- window(data.ts, start = c(1998, nTrain + 1), 
                   end = c(1998, nTrain + nValid))
```

### 1. Fit a regression model to the training data with a linear trend and additive seasonality 

```{r}
train.lm.trend.season <- tslm(train.ts ~ trend + season)
summary(train.lm.trend.season)
train.lm.trend.season.pred <- forecast(train.lm.trend.season, h = nValid)
```

### Create a plot to compare the fit to the training data. Do not show the validation data.

```{r}
plot(train.ts, ylab = "Victoria", xlab = "Time", bty = "l", 
     xlim = c(1998,2017), main = "")
lines(train.lm.trend.season.pred$fitted, lwd =2, col = "blue", lty = 3)
lines(train.lm.trend.season.pred$mean, lwd = 2, col = "blue", lty = 3)

```

### Create a plot to show the forecast and prediction interval.

```{r}
plot(train.lm.trend.season.pred, ylab = "Victoria", xlab = "Time", bty = "l", 
     xlim = c(1998,2017), main = "")
```

### 2. Fit an ARIMA model

#### What level of differencing do you need?
```{r}
ndiffs(train.ts)


nsdiffs(train.ts)

```
### So d = 0, D = 1

### Create an ACF and PACF plot on the differenced data.

```{r}

double.diff.train.ts <- (diff(train.ts, 4))
tsdisplay(double.diff.train.ts)
```

#### We can try any of p = 1 or q = 1 and P = 1 or Q = 1


### What AR and MA terms do you need?
```{r}
ARIMA1.fit <- Arima(train.ts, order = c(1,0,0), seasonal=c(1,1,0))
summary(ARIMA1.fit)
```


```{r}
ARIMA2.fit <- Arima(train.ts, order = c(0,0,1), seasonal=c(1,1,0))
summary(ARIMA2.fit)
```

```{r}
ARIMA3.fit <- Arima(train.ts, order = c(1,0,0), seasonal=c(0,1,1))
summary(ARIMA3.fit)
```

```{r}
ARIMA4.fit <- Arima(train.ts, order = c(0,0,1), seasonal=c(0,1,1))
summary(ARIMA4.fit)
```


### What ARIMA model would you recommend for this data? (If you are unsure then try several)


#### p = 1, d = 0, q = 0, P = 0, D = 1, Q = 1

### Fit your recommended ARIMA model

```{r}
man.ARIMA.fit <- Arima(train.ts, order = c(1,0,0), seasonal=c(0,1,1))
summary(man.ARIMA.fit)
```

```{r}
man.ARIMA.pred <- forecast(man.ARIMA.fit, h = nValid)
```

### Create a plot to compare the fit to the training data. Do not show the validation data.

```{r}
plot(train.ts, ylab = "Victoria", xlab = "Time", bty = "l", 
     xlim = c(1998,2017), main = "")
lines(man.ARIMA.pred$fitted, lwd = 2, col = "blue", lty = 3)
lines(man.ARIMA.pred$mean, lwd = 2, col = "blue", lty = 3)

```



### Create a plot to show the forecast and prediction interval.

```{r}
plot(man.ARIMA.pred, ylab = "Victoria", xlab = "Time", bty = "l", 
     xlim = c(1998,2017), main = "")
```



### 3.Use auto.arima() to fit an ARIMA (p, d, q) (P, D, Q) model to the training data

```{r}
auto.ARIMA.fit <- auto.arima(train.ts)
summary(auto.ARIMA.fit) 
```

```{r}
auto.ARIMA.pred <- forecast(auto.ARIMA.fit, h = nValid)

```


### Create a plot to compare the fit to the training data. Do not show the validation data.
```{r}
plot(train.ts, ylab = "Victoria", xlab = "Time", bty = "l", 
     xlim = c(1998,2017), main = "")
lines(auto.ARIMA.pred$fitted, lwd = 2, col = "blue", lty = 3)
lines(auto.ARIMA.pred$mean, lwd = 2, col = "blue", lty = 3)

```

### Create a plot to show the forecast and prediction interval.

```{r}
plot(auto.ARIMA.pred, ylab = "Victoria", xlab = "Time", bty = "l", 
     xlim = c(1998,2017), main = "")
```

### 4.Fitting an exponential smoothing model

```{r}
plot(train.ts, xlab = "Time", ylab = "Victoria", bty = "l")
```


### Looking at the training data:
### Is there a trend? What form does it take?
### Is there seasonality? What form does it take?

```{r}
ggseasonplot(train.ts) + labs(title="Seasonal Plot" )
```

#### From the above plots we can see that people tend to travel to Victoria mostly around Q1 and Q4.

```{r}
dec_data <- decompose(data.ts)
plot(dec_data)
```

#### We can see that there is additive seasonality with a Upward rising trend.
#### In other words, the seasonal magnitudes are the same every year and they add to the trend.



### What ETS() model would you recommend for this data? (If you are unsure then try several)

```{r}
ETS.model1 <- ets(train.ts, model = "AMA", restrict=FALSE) # Additive error, multiplicative trend, additive season
ETS.model1
```



```{r}
ETS.model2 <- ets(train.ts, model = "AMN", restrict=FALSE) # Additive error, multiplicative trend, no season
ETS.model2
```

```{r}
ETS.model3 <- ets(train.ts, model = "AAA", restrict=FALSE) # Additive error, additive trend, additive season
ETS.model3
```

```{r}
ETS.model4 <- ets(train.ts, model = "AAN", restrict=FALSE) # Additive error, additive trend, no season
ETS.model4
```

```{r}
ETS.model5 <- ets(train.ts, model = "MMA", restrict=FALSE) # Multiplicative error, multiplicative trend, additive season
ETS.model5
```

```{r}
ETS.model6 <- ets(train.ts, model = "MMN", restrict=FALSE) # Multiplicative error, multiplicative trend, no season
ETS.model6
```

```{r}
ETS.model7 <- ets(train.ts, model = "MAA", restrict=FALSE) # Multiplicative error, additive trend, additive season
ETS.model7
```

```{r}
ETS.model8 <- ets(train.ts, model = "MAN", restrict=FALSE) # Multiplicative error, additive trend, no season
ETS.model8

```

#### ETS Model 1 (A,M,A) is best with AIC = 1046.151

```{r}
ETS.model1.pred <- forecast(ETS.model1, h = nValid)
```


### Create a plot to compare the fit to the training data. Do not show the validation data.
```{r}
plot(train.ts, ylab = "Victoria", xlab = "Time", bty = "l", 
     xlim = c(1998,2017), main = "")
lines(ETS.model1.pred$fitted, lwd = 2, col = "blue", lty = 3)
lines(ETS.model1.pred$mean, lwd = 2, col = "blue", lty = 3)

```

### Create a plot to show the forecast and prediction interval.

```{r}
plot(ETS.model1.pred, ylab = "Victoria", xlab = "Time", bty = "l", 
     xlim = c(1998,2017), main = "")
```

### 5. Fit an ETS model allowing the algorithm to choose the structure for error, trend and seasonality from the training data (10 points)

```{r}
ETS.alg.model <- ets(train.ts, restrict = FALSE)
ETS.alg.model
```
#### The algorithmic model- ETS(M,N,M) with AIC = 1042.427

### Fit your recommended ets model.

```{r}
ETS.alg.model.pred <- forecast(ETS.alg.model, h = nValid)
```

### Create a plot to compare the fit to the training data. Do not show the validation data.
```{r}
plot(train.ts, ylab = "Victoria", xlab = "Time", bty = "l", 
     xlim = c(1998,2017), main = "")
lines(ETS.alg.model.pred$fitted, lwd = 2, col = "blue", lty = 3)
lines(ETS.alg.model.pred$mean, lwd = 2, col = "blue", lty = 3)

```

### Create a plot to show the forecast and prediction interval.

```{r}
plot(ETS.alg.model.pred, ylab = "Victoria", xlab = "Time", bty = "l", 
     xlim = c(1998,2017), main = "")
```

### 6. Assess the predictive accuracy of your five models in cross-validation (10 points)

```{r}
accuracy(train.lm.trend.season.pred, valid.ts)
accuracy(man.ARIMA.pred, valid.ts)
accuracy(auto.ARIMA.pred, valid.ts)
accuracy(ETS.model1.pred, valid.ts)
accuracy(ETS.alg.model.pred, valid.ts)

```
#### We could say ETS.model1 is best with least MAPE.

### 7.Which model would you recommend to the Australian tourism board for forecasting trips to the Victoria region  (20 points)

#### We would recommend the #4 (ETS.model1) Model to the Australian tourism board for forecasting trips to the Victoria region as it has best accuracy with considerable low error.


### Create a plot to compare the fit of your recommended model to the training and validation data.


```{r}
plot(train.ts, ylab = "Victoria", xlab = "Time", bty = "l", 
     xlim = c(1998,2017), main = "")
lines(ETS.model1.pred$fitted, lwd = 2, col = "blue", lty = 3)
lines(ETS.model1.pred$mean, lwd = 2, col = "blue", lty = 3)
lines(valid.ts, col = "grey20", lty = 3)
```

### Create a plot to show a 3-year forecast and prediction interval for your chosen model based on the full dataset.


```{r}
final_model=ets(data.ts, model = "AMA", restrict=FALSE)
final_model.pred <- forecast(final_model, h = 12)
```

```{r}
plot(final_model.pred, ylab = "Victoria", xlab = "Time", bty = "l", 
     xlim = c(1998,2020), main = "")
lines(final_model.pred$fitted, lwd = 2, col = "blue", lty = 3)
```



